### 2023-07-19 python_excel_ite.tex
fixed missing bib items, now editing

Spreadsheet based modeling has transformed the teaching of management science in business schools over the past twenty plus years. The late 1990's saw first editions of game changing textbooks \cite{winstonPracticalManagementScience2018,ragsdaleSpreadsheetModelingDecision2017} focused on the use of spreadsheets in teaching modeling and management science.  A series of conferences focused on teaching management science (TMS) with spreadsheets started at Dartmouth in 1998 and several presentations from that conference made up the bulk of papers in the first two issues of the INFORMS Transactions on Education journal. The next few decades saw numerous textbooks \cite{BusinessAnalyticsArt, BusinessAnalytics4th} and teaching papers \cite{INFORMSTransactionsEducation2014} that prominently featured spreadsheet based modeling.

@@

Python has continued to make inroads to analytical territory that was formerly dominated by spreadsheets. The widely used Python data analysis library, pandas, was created by Wes McKinney while working as a quantitative financial analyst \cite{PythonDataAnalysis}. Python is widely used for data wrangling, visualization and predictive modeling using various flavors of regression, tree based models and neural networks. You can get a very good sense of this by visiting the very popular \href{https://www.reddit.com/r/datascience/}{r/datascience subreddit on Reddit} -- a well known online forum. But, it got me to thinking. What about those things for which Excel is particularly well suited such as building formula based models and doing sensitivity analysis on these models? What would those look like in Python? The importance of modeling, articulated so well in the pages of this journal \cite{powellTeachingModelingManagement2001}, has not abated. What role does Python have to play in supporting the kind of modeling taught for years in our business schools?

In 2021, I published a \href{https://bitsofanalytics.org/#category=excel}{series of blog posts} that explored this question and formed the basis of the teaching module described in this paper. During the development of these posts, I came across the \href{https://quantecon.org/}{Quant Econ: open source code for economic modeling} project which \quote{is a nonprofit organization dedicated to development and documentation of open source computational tools for economics, econometrics, and decision making}. This is a mature project, run by renowned economists, with significant support. They have developed and released numerous high quality Jupyter notebooks exploring a range of computational economic and financial modeling topics. Here was an example of the Python for business usage I had been seeking. With this as inspiration, I decided to turn my blog post series into a teaching module called Excel with Python (EwP).

@@

Recently there has been more recognition of the need for data scientists to learn basic software engineering principes \cite{SoftwareEngineeringData,SoftwareEngineeringDataa,BuildingReproducibleAnalytical}.  The image of newly minted data science graduates developing spaghetti coded Jupyter notebooks not nearly ready for being put into production has reached \href{https://www.reddit.com/r/ProgrammerHumor/comments/y2xe3n/like_every_time_ever_when_the_devops_engineer/}{meme like status}. I hoped to address some of these issues in the EwP module. Sneakily, students would be exposed to some simple software engineering concepts during the process of building and analyzing familiar models. This is really no different than the integration of advanced Excel functions and techniques with spreadsheet modeling tutorials. Just as spreadsheet modeling students were able to easily wow their coworkers and managers with INDEX-MATCH and intermediate Pivot Table skills, I hoped that my students ability to properly document a Python function and use git and GitHub would give them a bit more professional credibility.

@@

At our institution, The EwP module is part of a business course entitled ``Advanced Analytics with Python'' (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900}{AAP}). The students taking this course have already taken the PCDA course within which they learned fundamental Python programming within the context of data analytics over seven weeks. The PCDA course ends with intro to the \href{https://scikit-learn.org/stable/index.html}{scikit-learn library} for doing predictive modeling in Python and the AAP course begins where the PCDA course ends. The EwP module could also be used as part of a semester long Python based analytics course. Most AAP students have also already taken "Business Analytics" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460}{BA}) which is a spreadsheet based introduction to business analytics. We do not strictly require the spreadsheet based course be taken before the PCDA or AAP courses as that course draws students from outside of our business programs who are strictly interested in programmatic tools such as Python. The BA and PCDA courses have historically been offered in face to face mode in a computer teaching lab. Since 2020, these two courses have been also offered as online, asynchronous courses. The newer AAP course has been offered since 2021 in an online asynchronous mode. All three courses have \href{http://www.sba.oakland.edu/Faculty/isken/teaching.html}{extensive course websites} that are publicly accessible, including access to all of the video content and supporting files.

@@

All of the Python modules in the PCDA course as well as the entire AAP course are taught using a combination of Jupyter notebooks, Python scripts, and videos walking through the notebooks. Jupyter notebooks, part of \href{https://jupyter.org/}{Project Jupyter}, are an example of \textit{literate computing} \cite{Jupyter_proposal} and allow one to create educational narratives that support iterative exploration and development. The ability to weave standard explanatory language with formal programming languages has made notebook computing a defacto standard in the data science education community \cite{perkel2018jupyter}. Jupyter notebooks can also include equations written in \LaTeX, data visualizations, and other multi-media. The technology is free, open source, runs on most computing platforms, and has a huge user community that has created a large number of resources for students, faculty and industry practitioners.

Central to Jupyter notebooks is a mix of \textit{code cells} and \textit{markdown cells}. Code cells contain executable code while markdown cells contain plain text using \href{URL}{markdown} for styling text. For example, Figure \ref{cells_edit_model} shows two markdown cells separated by a code cell as they look when in editing mode. After running the cells, they appear as show in Figure \ref{cells_after_running}. Mixing code and markdown cells makes it quite easy to create interactive tutorials in which students can not only run pre-written code, but can also add their own code -- all with the aid of supporting explanatory text. All of the blog posts upon which EwP is based were written in Jupyter notebooks. There are countless introductory tutorials for Jupyter notebooks, but the official Project Jupyter Documentations \cite{ProjectJupyterDocumentation} is a very good place to start if you are new to notebook computing.

@@

There are two overarching learning objectives for this \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900/mod3a_whatif.html}{first EwP submodule}. Students learn how Python might be used to do something they are quite familiar with doing in a spreadsheet -- model building and sensitivity analysis. At the same time, students encounter, as needed, more advanced Python programming concepts and techniques. In this way, the computational problem to be solved drives the introduction of more advanced Python concepts and techniques instead of such things being presented in a vacuum. I make it a point to show that there are often multiple possible approaches to a computational problem by being willing to abandon one approach and trying another. Emphasis is given to the lessons learned from each path explored. Every notebook is accompanied by screencasts that walk the students through the material and act as a guided tour with challenges to overcome along the way. The screenshot in \ref{fig:activities} shows the portion of the course web page that provides the high level roadmap through the notebooks and associated screencasts.

@@

At this point we discuss a fundamental difference between the computing model of Excel and that of Jupyter notebooks -- while spreadsheets respond automatically to changes, Jupyter notebooks require rerunning all code cells that include or follow the changed code. This difference in behavior between Jupyter notebooks and spreadsheets might be one of the most important concepts for students to internalize. A well known challenge associated with computational notebooks is the ease of running code cells out of order and introducing inadvertent errors. This point along with strategies to avoid it are discussed during the first week of the course which focuses on \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900/jupyter_notebooks.html}{notebook computing best practices as well as a historical look} at computational notebooks.
\subsubsection{Sensitivity analysis}

Since order quantity is the key decision variable, we might want to see how profit changes for different order quantities. In Excel, the Data Table tool provides an easy way to do this and is a staple of spreadsheet based management science textbooks. In a Data Table, a range of order quantities can be used as row or column input and one or more output values can be computed. In Python, the vectorized nature of the \href{https://numpy.org/}{NumPy} library \cite{NumPy}provides a simple way to accomplish the same thing. A range of order quantities is created as a NumPy array that can be used directly to compute vectors of all of the intermediate and final output variables. Much like Data Tables, vectorized computations avoid having to explicitly iterate, or loop, through a collection of input values.

@@

A few design dilemmas are posed and some non-working approaches illustrated as we try to decide on how to implement an n-way data table function for this new object oriented model. These meta reflections are intentionally included to better mimic actual software development and reinforce the idea that software development is much more than simply writing code to implement a perfectly thought out design. We forge ahead with some preliminary ideas and are confronted with an intermediate problem related to generating a list of dictionaries that represent combinations of inputs, or \textit{scenarios}, to evaluate. This provides a great opportunity to remind students of one of the great strengths of the open source software ecosystem -- we can leverage work done in other packages and even look at actual source code to see how something was implemented. In this case, the well known \href{https://scikit-learn.org/stable/}{scikit-learn} package \cite{ScikitlearnMachineLearning} has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely. This example also reminds the students of the value of being able to make sense of API documentation and of object oriented code written by others.

@@

Before trying to build our \code{goal\_seek} function we take a brief detour into basic root finding algorithms through packages like \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html#root-finding}{SciPy} \cite{SciPy} as well as in \href{https://github.com/patrickwalls/mathematical-python/}{open source notebooks and scripts} that are widely available in the Python data science ecosystem \cite{wallsMathematicalPython2023}. We compare the output of Excel's Goal Seek with different starting values to the values obtained by different root finding algorithms implemented in Python. The importance of starting values and also the range of different root finding algorithms available is often surprising to business students who have spent most of their analytical life in Excel.

@@

In the Excel modeling world, Monte-Carlo simulation can be done without add-ins, but packages like @Risk \cite{RISKRiskAnalysis} can make the process much easier. In particular, random variable generation for a wide range of probability distributions is facilitated by add-ins like @Risk. In Python, the NumPy package provide random variate generation for a large number of distributions and similarly, SciPy provides functions for computing distribution related quantities for many probability distributions. Much like we did when creating the \code{data\_table} function, we start with adding uncertainty to a single input variable in the bookstore model and rely on NumPy's vectorized computing capabilities to generate the profit associated with each realization of the random input. Then we can do standard statistical analysis of the simulation output using Python packages such as pandas and SciPy. In spreadsheet based simulation textbooks, the notion of the ``Flaw of Averages'' popularized by Sam Savage \cite{savageFlawAveragesWhy2012} is often illustrated. We do the same with our Python based model, showing how replacing random demand by its mean can result in a wildly optimistic estimate of the mean profit found through analyzing the simulation output. Much like doing simulation in Excel without add-ins, using Python makes transparent the quite simple and brute force nature of Monte-Carlo simulation.

@@

This is followed by a well commented first attempt at a \code{simulate} function. By this point, students are well equipped to understand the code as it builds on basic Python concepts and more advanced concepts that we have already covered in the EwP module. The rest of the notebook is spent trying out the \code{simulate} function and ends with creating a grouped boxplot that many of them have likely created before with @Risk along with a faceted plot of histograms of profit created with the popular \href{https://seaborn.pydata.org/}{Seaborn} package.

\begin{figure}[!ht]
	\centering
	\begin{minipage}{.80\textwidth}
		\includegraphics[width=1.0\textwidth]{images/boxplots.png}
	\end{minipage}
	\caption{Profit boxplots by order quantity}
	\label{fig:2way_dt}
\end{figure}

@@

The first notebook in this submodule starts by explaining what Python packages are and their role in sharing code. We then revisit a concept from earlier in the course -- creating a good project folder structure. For this we use what is known as a \href{https://cookiecutter.readthedocs.io/en/stable/}{cookiecutter} that can automatically generate a project folder structure and key files from the answers to a few prompts \cite{CookiecutterBetterProject}. There are a few different folder layouts used in Python projects and there is a bit of discussion on the folder structure we will use and resources given for exploring this issue further. Students are reminded of the importance of version control and we initialize a git repository for our budding package. The basics of version control were already covered in the first week of the course. My experience is that most students have little exposure to software project management concepts such as project folder structures and version control. This can lead to a jumble of files with incoherent names used as a proxy for version control -- \code{whatif\_v1.py}, \code{whatif\_v2.py}, \code{whatif\_final.py}, \code{whatif\_finalfinal.py}, and so on.

@@

This module has been included in my Advanced Analytics with Python course for the past three years. The course is offered in a seven-week accelerated format during the summer in an online and asynchronous mode. The EwP module is covered over the course of two weeks and is the subject of one of the major homework assignments for the semester. The first part of the assignment involves creating a basic model, similar in complexity to the Bookstore Model, and doing sensitivity analysis, goal seeking and simulation with the model. It requires the students to import and use the \code{whatif.py} package we created as part of the module. The second part of the assignment is based on the last submodule and requires  students to complete a task involving creation of a multi-sheet Excel workbook from a set of csv files using Python. Then, using the openpyxl library, they have to create and add formulas to the workbook and format them properly. The ranges used in the formulas vary by sheet and require determination of the number of rows in each sheet of data and use that information to construct the appropriate formula. An example assignment is available from \href{some.url}{here}. The module has been well received and several of the working students have reported back to me that they have found use cases for these techniques in their day to day analytical activities.

%TODO - how to link to assignment

Rumors have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA within their MS Office suite of packages \cite{MicrosoftConsidersAdding}. While this has not yet materialized, there continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations \cite{PythonpptxPythonpptx21}. Given that the ubiquity of both Excel and Powerpoint in business education and practice, inclusion of this material in a business school Python based analytics course seems quite appropriate. Students can learn more advanced and extremely useful Python concepts and techniques within a context with which they are intimately familiar. Manipulation of MS Office documents is directly relevant to the job content of many of the students. This teaching module is a natural complement to the commonly taught spreadsheet based analytics and introductory data science courses found in many business schools and provides an interesting way to teach and learn Python for analytics.

%TODO - Fix all of the bib entries with no authors. It's mostly websites.

>>>>

Spreadsheet based modeling has transformed the teaching of management science in business schools over the past twenty plus years. The late 1990's saw first editions of game changing textbooks \cite{winstonPracticalManagementScience2018,ragsdaleSpreadsheetModelingDecision2017} focused on the use of spreadsheets in teaching modeling and management science.  A series of conferences focused on teaching management science (TMS) with spreadsheets started at Dartmouth in 1998 and several presentations from that conference made up the bulk of papers in the first two issues of the INFORMS Transactions on Education journal. The next few decades saw numerous textbooks \cite{powellstepheng.BusinessAnalyticsArt, camm2020business} and teaching papers \cite{INFORMSTransactionsEducation2014} that prominently featured spreadsheet based modeling.

@@

Python has continued to make inroads to analytical territory that was formerly dominated by spreadsheets. The widely used Python data analysis library, pandas, was created by Wes McKinney while working as a quantitative financial analyst \cite{mckinney2022python}. Python is widely used for data wrangling, visualization and predictive modeling using various flavors of regression, tree based models and neural networks. You can get a very good sense of this by visiting the very popular \href{https://www.reddit.com/r/datascience/}{r/datascience subreddit on Reddit} -- a well known online forum. But, it got me to thinking. What about those things for which Excel is particularly well suited such as building formula based models and doing sensitivity analysis on these models? What would those look like in Python? The importance of modeling, articulated so well in the pages of this journal \cite{powellTeachingModelingManagement2001}, has not abated. What role does Python have to play in supporting the kind of modeling taught for years in our business schools?

In 2021, I published a \href{https://bitsofanalytics.org/#category=excel}{series of blog posts} that explored this question and formed the basis of the teaching module described in this paper. During the development of these posts, I came across the \href{https://quantecon.org/}{Quant Econ: open source code for economic modeling} project which ``is a nonprofit organization dedicated to development and documentation of open source computational tools for economics, econometrics, and decision making''. This is a mature project, run by renowned economists, with significant support. They have developed and released numerous high quality Jupyter notebooks exploring a range of computational economic and financial modeling topics. Here was an example of the Python for business usage I had been seeking. With this as inspiration, I decided to turn my blog post series into a teaching module called Excel with Python (EwP).

@@

Recently there has been more recognition of the need for data scientists to learn basic software engineering principes \cite{nelsoncatherineSoftwareEngineeringData2024,treadwayandrewSoftwareEngineeringData2023,rodriguesBuildingReproducibleAnalytical2023}.  The image of newly minted data science graduates developing spaghetti coded Jupyter notebooks not nearly ready for being put into production has reached \href{https://www.reddit.com/r/ProgrammerHumor/comments/y2xe3n/like_every_time_ever_when_the_devops_engineer/}{meme like status}. I hoped to address some of these issues in the EwP module. Sneakily, students would be exposed to some simple software engineering concepts during the process of building and analyzing familiar models. This is really no different than the integration of advanced Excel functions and techniques with spreadsheet modeling tutorials. Just as spreadsheet modeling students were able to easily wow their coworkers and managers with INDEX-MATCH and intermediate Pivot Table skills, I hoped that my students ability to properly document a Python function and use git and GitHub would give them a bit more professional credibility.

@@

At our institution, The EwP module is part of a business course entitled  \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900}{Advanced Analytics with Python} (AAP). The students taking this course have already taken the PCDA course within which they learned fundamental Python programming within the context of data analytics over seven weeks. The PCDA course ends with intro to the \href{https://scikit-learn.org/stable/index.html}{scikit-learn library} for doing predictive modeling in Python and the AAP course begins where the PCDA course ends. The EwP module could also be used as part of a semester long Python based analytics course. Most AAP students have also already taken   \href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460}{Business Analytics} (BA) which is a spreadsheet based introduction to business analytics. We do not strictly require the spreadsheet based course be taken before the PCDA or AAP courses as that course draws students from outside of our business programs who are strictly interested in programmatic tools such as Python. The BA and PCDA courses have historically been offered in face to face mode in a computer teaching lab. Since 2020, these two courses have been also offered as online, asynchronous courses. The newer AAP course has been offered since 2021 in an online asynchronous mode. All three courses have \href{http://www.sba.oakland.edu/Faculty/isken/teaching.html}{extensive course websites} that are publicly accessible, including access to all of the video content and supporting files.

@@

All of the Python modules in the PCDA course as well as the entire AAP course are taught using a combination of Jupyter notebooks, Python scripts, and videos walking through the notebooks. Jupyter notebooks, part of \href{https://jupyter.org/}{Project Jupyter}, are an example of \textit{literate computing} \cite{perez2015project} and allow one to create educational narratives that support iterative exploration and development. The ability to weave standard explanatory language with formal programming languages has made notebook computing a defacto standard in the data science education community \cite{perkel2018jupyter}. Jupyter notebooks can also include equations written in \LaTeX, data visualizations, and other multi-media. The technology is free, open source, runs on most computing platforms, and has a huge user community that has created a large number of resources for students, faculty and industry practitioners.

Central to Jupyter notebooks is a mix of \textit{code cells} and \textit{markdown cells}. Code cells contain executable code while markdown cells contain plain text using \href{URL}{markdown} for styling text. For example, Figure \ref{fig:cells_edit_mode} shows two markdown cells separated by a code cell as they look when in editing mode. After running the cells, they appear as show in Figure \ref{fig:cells_after_running}. Mixing code and markdown cells makes it quite easy to create interactive tutorials in which students can not only run pre-written code, but can also add their own code -- all with the aid of supporting explanatory text. All of the blog posts upon which EwP is based were written in Jupyter notebooks. There are countless introductory tutorials for Jupyter notebooks, but the official Project Jupyter Documentations \cite{jupyterteamProjectJupyterDocumentation2015} is a very good place to start if you are new to notebook computing.

\begin{figure*}[!htbp]
	\centering
	\caption{Notebook cells in edit model}
	\includegraphics[scale=1.0]{images/cells_edit_mode.png}
	\label{fig:cells_edit_mode}
\end{figure*}

\begin{figure*}[!htbp]
	\centering
	\caption{Notebook cells after execution}
	\includegraphics[scale=1.0]{images/cells_after_running.png}
	\label{fig:cells_after_running}
\end{figure*}

@@

There are two overarching learning objectives for this \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900/mod3a_whatif.html}{first EwP submodule}. Students learn how Python might be used to do something they are quite familiar with doing in a spreadsheet -- model building and sensitivity analysis. At the same time, students encounter, as needed, more advanced Python programming concepts and techniques. In this way, the computational problem to be solved drives the introduction of more advanced Python concepts and techniques instead of such things being presented in a vacuum. I make it a point to show that there are often multiple possible approaches to a computational problem by being willing to abandon one approach and trying another. Emphasis is given to the lessons learned from each path explored. Every notebook is accompanied by screencasts that walk the students through the material and act as a guided tour with challenges to overcome along the way.

@@

At this point we discuss a fundamental difference between the computing model of Excel and that of Jupyter notebooks -- while spreadsheets respond automatically to changes, Jupyter notebooks require rerunning all code cells that include or follow the changed code. This difference in behavior between Jupyter notebooks and spreadsheets might be one of the most important concepts for students to internalize. A well known challenge associated with computational notebooks is the ease of running code cells out of order and introducing inadvertent errors. This point along with strategies to avoid it are discussed during the first week of the course which focuses on \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900/jupyter_notebooks.html}{notebook computing best practices as well as a historical look} at computational notebooks.

\subsubsection{Sensitivity analysis}

Since order quantity is the key decision variable, we might want to see how profit changes for different order quantities. In Excel, the Data Table tool provides an easy way to do this and is a staple of spreadsheet based management science textbooks. In a Data Table, a range of order quantities can be used as row or column input and one or more output values can be computed. In Python, the vectorized nature of the \href{https://numpy.org/}{NumPy} library \cite{harris2020array} provides a simple way to accomplish the same thing. A range of order quantities is created as a NumPy array that can be used directly to compute vectors of all of the intermediate and final output variables. Much like Data Tables, vectorized computations avoid having to explicitly iterate, or loop, through a collection of input values.

@@

A few design dilemmas are posed and some non-working approaches illustrated as we try to decide on how to implement an n-way data table function for this new object oriented model. These meta reflections are intentionally included to better mimic actual software development and reinforce the idea that software development is much more than simply writing code to implement a perfectly thought out design. We forge ahead with some preliminary ideas and are confronted with an intermediate problem related to generating a list of dictionaries that represent combinations of inputs, or \textit{scenarios}, to evaluate. This provides a great opportunity to remind students of one of the great strengths of the open source software ecosystem -- we can leverage work done in other packages and even look at actual source code to see how something was implemented. In this case, the well known \href{https://scikit-learn.org/stable/}{scikit-learn} package \cite{scikit-learn} has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely. This example also reminds the students of the value of being able to make sense of API documentation and of object oriented code written by others.

@@

Before trying to build our \code{goal\_seek} function we take a brief detour into basic root finding algorithms through packages like \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html#root-finding}{SciPy} \cite{2020SciPy-NMeth} as well as in \href{https://github.com/patrickwalls/mathematical-python/}{open source notebooks and scripts} that are widely available in the Python data science ecosystem \cite{wallsMathematicalPython2023}. We compare the output of Excel's Goal Seek with different starting values to the values obtained by different root finding algorithms implemented in Python. The importance of starting values and also the range of different root finding algorithms available is often surprising to business students who have spent most of their analytical life in Excel.

@@

In the Excel modeling world, Monte-Carlo simulation can be done without add-ins, but packages like @Risk \cite{palisadeRISKRiskAnalysis} can make the process much easier. In particular, random variable generation for a wide range of probability distributions is facilitated by add-ins like @Risk. In Python, the NumPy package provide random variate generation for a large number of distributions and similarly, SciPy provides functions for computing distribution related quantities for many probability distributions. Much like we did when creating the \code{data\_table} function, we start with adding uncertainty to a single input variable in the bookstore model and rely on NumPy's vectorized computing capabilities to generate the profit associated with each realization of the random input. Then we can do standard statistical analysis of the simulation output using Python packages such as pandas and SciPy. In spreadsheet based simulation textbooks, the notion of the ``Flaw of Averages'' popularized by Sam Savage \cite{savageFlawAveragesWhy2012} is often illustrated. We do the same with our Python based model, showing how replacing random demand by its mean can result in a wildly optimistic estimate of the mean profit found through analyzing the simulation output. Much like doing simulation in Excel without add-ins, using Python makes transparent the quite simple and brute force nature of Monte-Carlo simulation.

@@

This is followed by a well commented first attempt at a \code{simulate} function. By this point, students are well equipped to understand the code as it builds on basic Python concepts and more advanced concepts that we have already covered in the EwP module. The rest of the notebook is spent trying out the \code{simulate} function and ends with creating a grouped boxplot, see Figure \ref{fig:boxplot} that many of them have likely created before with @Risk along with a faceted plot of histograms of profit created with the popular \href{https://seaborn.pydata.org/}{Seaborn} package.

\begin{figure}[!ht]
	\centering
	\begin{minipage}{.80\textwidth}
		\includegraphics[width=1.0\textwidth]{images/boxplots.png}
	\end{minipage}
	\caption{Profit boxplots by order quantity}
	\label{fig:boxplot}
\end{figure}

@@

The first notebook in this submodule starts by explaining what Python packages are and their role in sharing code. We then revisit a concept from earlier in the course -- creating a good project folder structure. For this we use what is known as a \href{https://cookiecutter.readthedocs.io/en/stable/}{cookiecutter} that can automatically generate a project folder structure and key files from the answers to a few prompts \cite{audreyroyandcookiecuttercommunityCookiecutterBetterProject2012}. There are a few different folder layouts used in Python projects and there is a bit of discussion on the folder structure we will use and resources given for exploring this issue further. Students are reminded of the importance of version control and we initialize a git repository for our budding package. The basics of version control were already covered in the first week of the course. My experience is that most students have little exposure to software project management concepts such as project folder structures and version control. This can lead to a jumble of files with incoherent names used as a proxy for version control -- \code{whatif\_v1.py}, \code{whatif\_v2.py}, \code{whatif\_final.py}, \code{whatif\_finalfinal.py}, and so on.

@@

This module has been included in my Advanced Analytics with Python course for the past three years. The course is offered in a seven-week accelerated format during the summer in an online and asynchronous mode. The EwP module is covered over the course of two weeks and is the subject of one of the major homework assignments for the semester. The first part of the assignment involves creating a basic model, similar in complexity to the Bookstore Model, and doing sensitivity analysis, goal seeking and simulation with the model. It requires the students to import and use the \code{whatif.py} package we created as part of the module. The second part of the assignment is based on the last submodule and requires  students to complete a task involving creation of a multi-sheet Excel workbook from a set of csv files using Python. Then, using the openpyxl library, they have to create and add formulas to the workbook and format them properly. The ranges used in the formulas vary by sheet and require determination of the number of rows in each sheet of data and use that information to construct the appropriate formula. An example assignment is available from \href{https://drive.google.com/file/d/1prf8acZXZ4RoYBgQ8e9-eijA65uV7T-V/view?usp=sharing}{here}. The module has been well received and several of the working students have reported back to me that they have found use cases for these techniques in their day to day analytical activities.

Rumors have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA within their MS Office suite of packages \cite{cimpanucatalinMicrosoftConsidersAdding}. While this has not yet materialized, there continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations \cite{cannystevePythonpptx2013}. Given that the ubiquity of both Excel and Powerpoint in business education and practice, inclusion of this material in a business school Python based analytics course seems quite appropriate. Students can learn more advanced and extremely useful Python concepts and techniques within a context with which they are intimately familiar. Manipulation of MS Office documents is directly relevant to the job content of many of the students. This teaching module is a natural complement to the commonly taught spreadsheet based analytics and introductory data science courses found in many business schools and provides an interesting way to teach and learn Python for analytics.
